Interactive Filter and Display of Hillary Clinton's Emails: \\A Cautionary Tale of Metadata

### 2017-04-02 VAST_submission/vgtc_journal_latex/paper.tex
Introduced subsections, changed ten patches to nine

Following this approach, ten more patches can be found (in addition to that of Section~\ref{sect:emailGap}).  With the dates in hand, an internet search can be conducted to see what, if anything, might turn up to explain the patch.
\subsubsection{You have no mail \ldots}
\label{sect:youHaveNoMail}
The first, and largest, patch occurs at the very beginning of Clinton's tenure as Secretary of State.  It is  obvious even from the full time series of Figure~\ref{fig:VolumeAll}.  Essentially no email from Secretary Clinton, and little from anyone else appears until mid to late April 2009.

This is the time when Secretary Clinton's email habits first run afoul of the State Department as Clinton insists on using her personal Blackberry for all of her email communications.  The issue came to a head in February when National Security officials and intelligence specialists explain the risks to Cheryl Mills.  The officials were apparently not aware  of the private email server (e.g. see \cite{TakingRootWashPost}).  
In time order, the patches and what we learned from each Google search was as follows.

Three of the patches appear to be easily interpreted.   June 17 to June 23, 2009 coincides with Clinton's surgery for an elbow fracture after a fall \cite{ClintonFracture2009}.  A month later, from July 17 to July 23 another patch appears.  This coincides with Secretary Clinton's State visits to India and Thailand \cite{ForeignSched}.  Similarly, the absence of email from Clinton from October 11 to  October 13, 2009 corresponds to a State visit to Russia as shown in  Figure~\ref{fig:ClintonZerosOctober2009AndForeignSchedule} \cite{ForeignSched, visitRussia}.
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.95\linewidth]{ClintonZerosOctober2009AndForeignSchedule}
\caption{Oct 3 -- Oct 27, 2009:  Selected Emails are from Clinton; Scheduled foreign travel marked in blue.}
\label{fig:ClintonZerosOctober2009AndForeignSchedule}
\end{center}
\end{figure}

The next patch chronologically comes in 2011, from April 11 to April 15.   On the last two days of the patch, Secretary Clinton met with other NATO Foreign Ministers in Germany \cite{ForeignSched}.   Notably, in an April 10, 2011 email, Huma Abedin forwards Clinton the ``Stevens Update'' from Timmy Davis.  The email describes Ambassador Stevens's concerns about the situation in Ajdabiyah (less than 150 km from Benghazi) where he is considering leaving Benghazi \cite{StevensUpdate}.

From August 27 to August 29, and again in September 1, 2011 no emails are sent from Clinton.   No foreign travel appears on the schedule.  Notable events include ``Russian hackers'' tried to gain access to Clinton's email accounts through phishing \cite{russianHackers2}; a significant Clinton Foundation donor, Rajiv K. Fernando, was added by Cheryl Mills to the International Security Advisory Board  \cite{clintonDonorSecurity}; State Department executive secretary Stephen Mull is rebuffed by Huma Abedin when he suggests replacing Clinton's personal BlackBerry by a ``department-issued'' one \cite{earlyEmails}.

The next patch appears at the end of March, the beginning of April, 2012.  The exact dates were March 27 and 29, the range from March 31 to April 3, and then again on April 10.   Few, but not zero, emails were sent between these dates.  Google search does not turn up much other than the fact that Clinton Staffers seem to be vetting email \cite{tightReinRecords, staffVetEmail}, and that Clinton had received an email which in part contained classified material \cite{classifiedInPart}.

No emails are found sent from Secretary Clinton's email from August 27 to August 30 and September 1, 2, and 6, 2012, all just before the Benghazi attack.  Much has already been written about this time period.  Here we add that in a cable to ``SECSTATE'' dated August 8 (via Hanna Draper) and entitled ``The Guns of August: security in eastern Libya'' from Ambassador Stevens was the comment ``What we have seen are not random crimes of opportunity but rather targetted and discriminate attacks'' \cite{gunsOfAugust}.  According to the Benghazi Report,  during August the number of State Department security agents assigned to the embassy in Tripoli dropped from 34 to 6 \cite{BenghaziReport}.

The last patch appears shortly after the email gap from Section~\ref{sect:emailGap}.  Figure~\ref{fig:ZeroEmailsDec2012}
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.95\linewidth]{ZeroEmailsDec2012}
\caption{Dec 3 -- Jan 14, 2012:  Selected Emails are from Clinton; Scheduled foreign travel marked in blue.}
\label{fig:ZeroEmailsDec2012}
\end{center}
\end{figure}
shows two closely related gaps in December 2012 which, for the purposes of discourse only,  we are counting as only one of our ten patches.  Similar to June 2009, there is an easy explanation at hand.  Clinton had suffered a concussion ``\ldots in a fall brought on by an illness'' \cite{concussion} and was recovering at home by December 15.  She had `already been forced to cancel a planned trip to the Middle East and north Africa'' \cite{concussion} that week; this is not the foreign travel shown in Figure~\ref{fig:ZeroEmailsDec2012} \cite{ForeignSched}.

>>>>

Following this approach, nine %ten 
more patches can be found (in addition to that of Section~\ref{sect:emailGap}).  With the dates in hand, an internet search can be conducted to see what, if anything, might turn up to explain the patch.
%\subsubsection{You have no mail \ldots}
\subsubsection{Early days}
\label{sect:youHaveNoMail}
The first, and largest, patch occurs at the very beginning of Clinton's tenure as Secretary of State.  It is  obvious even from the full time series of Figure~\ref{fig:VolumeAll}.  Essentially no email from Secretary Clinton, and little from anyone else appears until mid to late April 2009.

This is the time when Secretary Clinton's email habits first run afoul of the State Department as Clinton insists on using her personal Blackberry for all of her email communications.  The issue came to a head in February when National Security officials and intelligence specialists explain the risks to Cheryl Mills.  The officials were apparently not aware  of the private email server (e.g. see \cite{TakingRootWashPost}).  
%In time order, the patches and what we learned from each Google search was as follows.

\subsubsection{Three patches in 2009}
Three of the patches appear to be easily interpreted.   June 17 to June 23, 2009 coincides with Clinton's surgery for an elbow fracture after a fall \cite{ClintonFracture2009}.  A month later, from July 17 to July 23 another patch appears.  This coincides with Secretary Clinton's State visits to India and Thailand \cite{ForeignSched}.  Similarly, the absence of email from Clinton from October 11 to  October 13, 2009 corresponds to a State visit to Russia as shown in  Figure~\ref{fig:ClintonZerosOctober2009AndForeignSchedule} \cite{ForeignSched, visitRussia}.
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.95\linewidth]{ClintonZerosOctober2009AndForeignSchedule}
\caption{Oct 3 -- Oct 27, 2009:  Selected Emails are from Clinton; Scheduled foreign travel marked in blue.}
\label{fig:ClintonZerosOctober2009AndForeignSchedule}
\end{center}
\end{figure}

\subsubsection{April 11 to April 15, 2011}
%The next patch chronologically comes in 2011, from April 11 to April 15.   
On the last two days of %the 
this patch, Secretary Clinton met with other NATO Foreign Ministers in Germany \cite{ForeignSched}.   Notably, in an April 10, 2011 email, Huma Abedin forwards Clinton the ``Stevens Update'' from Timmy Davis.  The email describes Ambassador Stevens's concerns about the situation in Ajdabiyah (less than 150 km from Benghazi) where he is considering leaving Benghazi \cite{StevensUpdate}.

\subsubsection{August 27 to August 29, September 1, 2011, 2011}
%From August 27 to August 29, and again in September 1, 2011 no emails are sent from Clinton.   
No foreign travel appears on the schedule.  Notable events include ``Russian hackers'' tried to gain access to Clinton's email accounts through phishing \cite{russianHackers2}; a significant Clinton Foundation donor, Rajiv K. Fernando, was added by Cheryl Mills to the International Security Advisory Board  \cite{clintonDonorSecurity}; State Department executive secretary Stephen Mull is rebuffed by Huma Abedin when he suggests replacing Clinton's personal BlackBerry by a ``department-issued'' one \cite{earlyEmails}.

\subsubsection{End of March,  beginning of April, 2012}
%The next patch appears at the end of March, the beginning of April, 2012.  
The exact dates were March 27 and 29, the range from March 31 to April 3, and then again on April 10.   Few, but not zero, emails were sent between these dates.  Google search does not turn up much other than the fact that Clinton Staffers seem to be vetting email \cite{tightReinRecords, staffVetEmail}, and that Clinton had received an email which in part contained classified material \cite{classifiedInPart}.

\subsubsection{End of August,  beginning of September, 2012}
No emails are found sent from Secretary Clinton's email from August 27 to August 30 and September 1, 2, and 6, 2012, all just before the Benghazi attack.  Much has already been written about this time period.  Here we add that in a cable to ``SECSTATE'' dated August 8 (via Hanna Draper) and entitled ``The Guns of August: security in eastern Libya'' from Ambassador Stevens was the comment ``What we have seen are not random crimes of opportunity but rather targetted and discriminate attacks'' \cite{gunsOfAugust}.  According to the Benghazi Report,  during August the number of State Department security agents assigned to the embassy in Tripoli dropped from 34 to 6 \cite{BenghaziReport}.

\subsubsection{December 2012}
The last patch appears 
shortly after the email gap from Section~\ref{sect:emailGap}.  Figure~\ref{fig:ZeroEmailsDec2012}
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.95\linewidth]{ZeroEmailsDec2012}
\caption{Dec 3 -- Jan 14, 2012:  Selected Emails are from Clinton; Scheduled foreign travel marked in blue.}
\label{fig:ZeroEmailsDec2012}
\end{center}
\end{figure}
shows two closely related gaps in December 2012 which, for the purposes of discourse only,  we are counting as only one of our 
%ten 
nine
patches.  Similar to June 2009, there is an easy explanation at hand.  Clinton had suffered a concussion ``\ldots in a fall brought on by an illness'' \cite{concussion} and was recovering at home by December 15.  She had `already been forced to cancel a planned trip to the Middle East and north Africa'' \cite{concussion} that week; this is not the foreign travel shown in Figure~\ref{fig:ZeroEmailsDec2012} \cite{ForeignSched}.
